\chapter{Site Administration}

On most Sites it is necessary to have an administration Department, header by a capable
administrator.

\section{Size of the administration Office}

To function properly the minimum size of the Administration Department is as follows:

\begin{table}
\begin{tabular}{ll}
\toprule
Designation    &Number\\
\midrule
Manager        &1\\
Admin Officer & 1\\
Time Keepers & 2\\
Driver            &1\\
\bottomrule
\end{tabular}
\caption{Minimum size for administration Site Department.}
\end{table}

The size of the Department will ultimtely be determined from the number of personnel working on the size
and can be increased or decreased during the life cycle of the Project. Given the Competitive pricing
this is one area where efficiency is important to keep costs down. 

\section{What does the Department do?}

For most personnel their encounter with the Company face is the Administartion Department. As such
it is important to keep in mind that people are highly demotivated when faced with inefficiencies of 
administration, especially during difficulties in the employess life, such as a death in the family.

So the first and most important function of the administration department is to act as an internal PRO
Department, helping to instill a culture of professionalism and efficiency. Most matters that relate
with personnel except for their technical work are handled by the Administartion Department.

\subsection*{Processing of Routine Documents}

The processing of daily routine stuff such as leave applications, resignations, demobilization and 
mobilization of personnel, time keeping, issuing of general memos, liaising with counter parts and follow up
with HO personnel.

\subsection*{Medical}

The department is repsonsible to make arrangments for the Medical treatment of employees and to keep
statistics in this respect.

\subsection*{Camps}

\subsection*{Transportation}

\subsection*{Disciplinary procedures}

\subsection*{Site Office Management}

If the Site is large enough, it will employ secretarial, office cleaning, photocopying and other personnel that
will come under the Management of the Administration Department.

\subsection{IT Department}

Id teh Site has a lot of computers the IT Technician or PC support Engineer, may also report
to the administartion department.

A back-up plan and disaster recovery will have to be planned by the Project Director and Heads of
Department.

\subsection{Visas, tickets and similar items}

\subsection{Payment of Salaries}

\subsection{Reporting}

Daily reports must be produced for attendance, deployment of personnel.

Weekly reports should provide a summary of the following:

\begin{tabular}{ll}
Man hours worked          &\\
Percent absenteeism       &\\
Total employees on site   &\\
Total employees on leave &\\
\end{tabular}

The actual format of the report will vary from Site to Site and will also vary if the planner is producing 
similar statistics.

Monthly reports will have a similar structure.

\subsection*{Document Control and Filing}

The section shall keep all documents as described earlier for Document Control. In particular the
following must be kept as a minimum:

\begin{tabular}{ll}
Employee files &\\
Weekly reports &\\
Personnel Movement &\\
Internal Memos &\\
Warnings &\\

\subsection*{Key performance Indicators}

Perhaps the best key performance indicator of the Department is that personnel trust that their paperworks have been processed. At City Center project for example having had to rely on an inefficeint HO it was not unknown for personnel to daily visit the Admin Office to enquire about their leave or visa status daily. 

\end{tabular}





















