\chapter{Engineering}

\newthought{Finalization} of Shop drawings for status 'A' and subsequent issue of these drawings \texttt{AS-BUILT} is going well for the Mechanical Drawings, whereas for the Electrical drawings it has been slowed down with additional changes and EIs. The Electrical reviews also are mostly delayed by the Engineer and in addition are returned with additional comments. This causes problems both for the finalization of drawings as well as the physical construction works.

We have now taken a strong approach to pressurize the Engineer as follows:

\begin{enumerate}[a.)]
\item Weekly letters are issued for general delays in returning these drawings.
\item All additional comments on drawings that constitute new work or alterations are queried via RFIs.
\item Any abnormal requests are responded to via letters.
\end{enumerate}

This increase in correspondence is having an effect and we have seen some effort to resolve the situation albeit too late and too little. In addition, it appears that the only way the Engineer can complete the design, is when drawings are in Status-A. HOK are very reluctant to issue Engineer's Instructions for areas where drawings are completed or the physical works are completed.

It is a pity that we have reduced personnel in the Engineering Office, as certainly reaching completion with drawings completed, makes life easier. It also saves a lot of trouble during commissioning.

\section{Engineering works for RFIs}

These are currently approximately 50-100 RFIs backlog, ditto for Engineer's instructions. There is also still considerable work in closing comments for problematic submittals.