\newcommand{\elvon}{$30^{th}$\/Jan 2011}
\chapter{ELV Systems}

\newthought{Completion of the majority of the ELV Systems} are expected to be delayed due to the issue of  EI 500. The EI affects most of the areas of the development and has seriously increased the amount work related to the ELV Systems. The EI does not provide a detail design and a substantial amount of work is expected to be required to revise drawings and clear commercial issues with  suppliers and sub-contractors. 
\begin{margintable}
	    \begin{tabular}{p{2.5cm}l}
	      \toprule
	      System    & Milestone \\
	      \midrule
                 Fire Alarm     &  \elvon  \\   
                 Security, Access Control and CCTV  & \elvon\\
	     PA and Backgorund Music System     &  \elvon  \\
                IPTV and Sattelite System &\elvon\\
     	      \bottomrule
	    \end{tabular}
           \caption{ELV Expected Completion Dates}
           \label{tbl:LVkeydates}
  \end{margintable}


We are not sure at the total impact of this EI at this stage but we expect these works to go post December, assuming a reasonable amount of time for procurement and engineering. Table \ref{tbl:LVkeydates} shows the expected completion dates based on these assumptions. We have also assumed that no further changes will be effected.  \sidenote{The impact of ELV systems being delayed is not localized to these systems. A number of services are integrated with the Fire Alarm etc., and these can only be tested and commissioned once the ELV Systems are operational. Civil Defence inspections will have to be postponed - until after these integration  tests are completed and hence the Operational License for the Buildings is expected to be delayed.}

 
 

\section{Current Status}

Most of these systems are currently at an advance stage in the Towers and Podia. Containment is progressing with Tower floors and was expected to be completed at around the end of August. However, we have not as yet ascertained the full impact of EI 500 and will revisit the issue in later revisions of this report.

\section{Constraints}

Current constraints are in Design, Procurement and Commercial Departments. 

\ramadaneffect
 

\section{Work Organization}

The person responsible for these works are the respective Tower and Podium Project Managers. 

We expect to re-organise these works once we are clearer as to when amended Shop Drawings can be issued.   

\section{Sequence for Switch-on}

In order to enable progressive commissioning early switch-on will be targetted. We will enter discussions with relevant Suppliers to propose interim solutions. However, we do not expect to be able to substantially bring dates forward and the dates shown in Table \ref{tbl:LVkeydates} can be considered as optimistic.










