
\chapter{Smoke Exhaust}


\newthought{The Smoke Extract Systems} can be completed and possibly commissioned by the \deadline. These systems are located at the top of all three Towers. They are expected to be completed in tandem with the Towers. However, work still remains to be completed on the top floors and motorized dampers are still to be delivered and a local supplier has still not been approved. There remains detailing works for the controls serving these systems. 

           \begin{margintable} 
	    \begin{tabular}{lcl}
	      \toprule
	      System    & Milestone \\
	      \midrule
                 Rotana      &  \ecolon  \\   
                 Shangri-la  & \ecolon\\
	     Merweb     &  \ecolon  \\
     	      \bottomrule
	    \end{tabular}
           \caption{Smoke Exhaust Key Dates}
           \label{tbl:SEkeydates}
            \end{margintable}
 

\section{Current Status}

The vertical and horizontal ductwork is progressing well in tandem with the floors. The Plantroom works have still to be completed. Electrical works and BMS works are still outstanding.

\section{Constraints}

Current constraints are human resources for ductwork installation (especially for Merweb). Delivery of ductwork (needs to be fire treated and it takes two trips, before arriving to Site). Design works for the plant rooms is still to be closed, although drawings are available.

\ramadaneffect
 

\section{Work Organization}

The persons  \sidenote{Although allocating Specon Senior Supervision to these areas is depleting our strength, by giving priority to the Smoke Extract and Staircase Pressurization systems to be commissioned by the \deadline\  places HEE and Specon in a strong position by the \deadline .}responsible for these works are the respective Tower Project Managers. The Senior Supervisors listed below are expected to take charge of the remaining works and complete the top floor plant rooms. The respective Tower Mechanical Engineers are expected to verify all materials, ensure ductwork is ordered early and to make arrangements for all inspections.

   \begin{table}[htbp] 
	    \begin{tabular}{l p{4cm}l}
	      \toprule
	     Item & Responsibility  \\
	      \midrule
                  1 &Rotana & \lambros \\
                  2 &Shangrila &\sotiris \\
	      3  &Merweb  & \sotiris  \\
                 4  &All electrical &\lukas\\
                   \bottomrule
   \end{tabular}
 \caption{Staircase Pressurization responsible persons}
\end{table}


\section{Sequence for Switch-on}

In order to enable progressive commissioning early switch-on will be targetted. It is expected that these fans can be activated shortly after power is provided to the Towers. The BMS controls will only be activated once all DDC panels are installed and the BMS nears completion.










