\chapter{Towers}\label{sec:towers1}\index{towers}
\index{Towers! progress}

\newthought{The Hotel Towers can be completed by October 30th}\sidenote{Completion meaning, substantial completion and preliminary operation and testing where possible}. Problems have mostly been resolved, although there are still issues with ID,  EIs and inspections.  The Merweb Tower - which always - lagged behind the other Towers continues to do so. This Tower will need a concerted effort and additional personnel by the subcontractor to meet the deadline. We have factored this into our completion expectations.

Currently Tower focus is on closing walls and ceilings without at the same time neglecting other important issues. For example we are now busy with completing all the busducts in order to enable full power-on as soon as Kahraama power becomes available.

Ceiling closures, are progressing better in both Rotana and Shangri-la and we are convinced that Merweb canl pick-up speed, as JEM personnel become more familiar with the works and re-organize as advised.

Specon is assisting all subcontractors, through tradesmen and supervision  in order to secure the deadlines for completion.

\section{Constraints}

Materials are less of an issue now than they were a month ago. However, the effects of earlier delays in deliveries has taken its toll. Ductwork is still problematic but the shortages have been alleviated to a large degree, by the decision to set-up of a Specon manufacturing workshop on Site.

\section{Current Status}
The status of the works are summarized in the Tables that follow. We expect planned dates to vary slightly from those given to accomodate unforeseen problems. Overall, though we do expect the final completion dates as stated in the tables to be close to those indicated.








